Designing Qusecs Optimally: Quasi-uniform or Self-similar Layered Materials as Recursively Decomposed Solutions of Geometric Constraint Systems

### 2014-12-17 doc/GMP2015/s02_preliminaries.tex
Updated preliminary. Added not to recombination. Modified pinned line-incidence according to prelimary and updated DRP section.

\subsection{Basics}
%TODO!

In this paper, a {\em geometric constraint system} is a multivariate polynomial 
(usually bilinear or quadratic) system $G(x,\delta)=0$, representing constraints with parameters $\delta$ 
between geometric primitives  in $\mathbb{R}^2$ represented collectively $x\in \mathbb{R}^n$.   
%
When the type of constraint (system) is fixed, the system is simply represented as $(G,\delta)$, where 
$G$ is the underlying constraint (hyper)graph $G = (V,E)$ 
with the vertices $V$ representing the geometric primitives in 
$\mathbb{R}^2$ and (hyper)edges
$E$ representing the constraints, each with an associated 
parameter $\delta$.
%
For example, a {\em bar-joint system or linkage} $(G,\delta)$, is a graph
$G=(V,E)$  
with fixed length bars as edges, i.e. $\delta: E \rightarrow \mathbb{R}$;
this represents the distance constraint system $\| x_u -x_v \|_2 =
\delta_{u,v}$ for  $(u,v) \in E$, where $x_u \in \mathbb{R}^2$ represents the coordinates of
$u\in V$.

@@

Some of these generic constraint system or 
(hyper)graph properties are the following ({\bf Note:} we refer to these as
properties of the constraint system or as properties of the underlying
(hyper)graph interchangeably).
%
A constraint (hyper)graph $G$ is {\em independent} if its appropriate rigidity matrix of indeterminates  
has independent rows
(i.e, the determinant of some square submatrix is not
identically zero). 
It is {\em rigid} if the number of independent rows of the rigidity matrix is
at least the number of columns less the number of rigid body motions, 
which is 3 for distance constraint systems. 
It is {\em isostatic (minimally rigid, well-constrained)} 
if it is both rigid and independent. 
%if the number of generically independent rows or the rank of the appropriate rigidity matrix is maximal.
%For example, the maximal rank of a bar-joint rigidity matrix is $2|V| - 3$, 
%where $3$ is the number of rotational and translational degrees-of-freedom of a rigid body in $\mathbb{R}^2$.
%The graph $G$ is {\em rigid} if there exists some spanning subgraph $S\subseteq G$ such that $S$ is well-constrained. 
It is {\em flexible} if it is not rigid, {\em under-constrained} if it is independent and not rigid, or {\em over-constrained} if it is not independent.

In this paper, a {\em qusec} is any {\em independent} geometric 
constraint system of one of the 3
types mentioned above.

The 2-dimensional {\em rigidity matroid} of a constraint (hyper)graph $G = (V,E)$ is a linear matroid  on $E$, 
where a subset of edges $E' \subseteq E$ is {\em independent} in the matroid, 
if the set of corresponding rows in the appropriate rigidity matrix of indeterminates are linearly independent.

There are various results on combinatorial characterization of independence and rigidity
for different types of  rigidity matroid. 
For bar-joint rigidity matroid, 
the famous Laman's theorem~\cite{???} states that the underlying graph
is minimally rigid if and only if $|E| = 2|V|-3$ and $|E'| \le 2|V'|-3$ for every induced subgraph with at least 2 vertices. 
The result by Lovasz and Yemini~\cite{???} shows that all 6-vertex-connected graphs are rigid in the plane. 
For bar-body rigidity matroid,
Tay~\cite{???} proved that
the underlying multigraph is minimally rigid if and only if it can be decomposed as $3$ edge disjoint spanning trees. 
White and Whiteley~\cite{???} gave the same characterization using a different technique to study the algebraic - geomtric conditions called pure condition.
More generally, 
Lee, Streinu and Theran~\cite{???} defined the {\em $(k,l)$-sparsity matroid}.
A hypergraph $G$ is called {\em $(k,l)$-sparse} if $|E'| \le k|V'| - l$ for any induced subgraph $(V',E')$ with at least 2 vertices, 
and {\em $(k,l)$-tight} if is $(k,l)$-sparse and $|E| = k|V| - l$. 
In general, if you have a $d$-uniform hypergraph, a $(k,l)$-sparsity
condition is matroidal as long as $l \le dk-1$.

For bar-joint graphs,  
a vector  in the right null space of its rigidity matrix
is called a {\em flex vector}. 
It has 2 entries of indeterminates per vertex and represents the internal motions of the system
(i.e.\ modulo rigid body motions).
For rigid graphs, all flex vectors are  identically zero. 
A vector in the left null space of the rigidity matrix
is called a {\em stress vector}, with one entry of indeterminate per edge.
For independent graphs, all stress vectors are identically zero.

%--A generalized version of Asimow Roth (implicit fn thm)
%generic rigidity equivalent to infinitesimal rigidity (easy
%when the latter def is based on indeterminates)

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

%XXXXXXXXXXXXXXX BASICS until combinatorial characterization of 
%independence and rigidity XXXXXXXXXXX

NOTE: geometric constraint systems can have inequalities (the parameters
in $\delta$ being inexact, small intervals of values).

XXXXXXXXXsomething has to be said 
along with asimow roth (infinitesimal rigidity 
implies rigidity)
when the rigidity matrix has appropriate rank
the realizations or solutions of the 
of the corresponding constraint system 
are generically isolated and zero-dimensional 
(if the constraint system has equalities and exact 
values for the parameters $\delta$.
If the constraint system has inequalities, i.e, 
if the $\delta$ lie in an interval,
this claim is approximate: 
the solutions are isolated small, full-dimensional  connected components.

%XXXXXXXX Also in the recombination section, it should be noted
%that the HUGE advantage of the convex Cayley method
%is that it is completely unaffected by having the $\delta$ be intervals
%cite my paper with Heping!!!

%XXXXXXXXXX CUrrently there are open problems strewn throughout the paper. 
%Should they be collected to the conclusion?
%Or at least clearly labeled as open problems?
%XXXXXXXXX

\noindent{\bf Note:}
In the remainder of this section and the next two sections 
\ref{sec:DRP} and \ref{sec:recomb} 
we only consider bar-joint qusecs and graphs
Relevant formal analogies  for the other 2 types of qusecs and (hyper)graphs 
are given in the subsequent 2 sections on Applications.

XXXNow comes the definition of stress vector and flex vector. 
NOTE: something has to be said about what flexes and stresses mean

An {\em external stress} $t$ is a $1\times 2|V|$ vector  (of indeterminates)
which specifies a 2D vector acting at each vertex
such that the stress balance equation $sR = t$ holds, where $s$ is a vector
of internal stresses, one for each bar, and $R$ is the rigidity matrix.

\subsection*{Decomposition-Recombination (DR) Plans}

@@

A \textbf{trivial} graph is   a single vertex.
\begin{definition}\label{def:drp}
    The \textbf{decomposition-recombination \mbox{(DR-)} plan} of graph $G$,
    \cite{XX}
    $DRP(G)$, is defined as a forest that has the following properties:
    \begin{enumerate}
        \item Each  node represents/contains/is  a rigid vertex-induced subgraph of $G$
        %\item For a node, $C$, that is  a non-trivial graph, its $N$
        %    children, $C_1, \ldots, C_N$, are rigid vertex-maximal proper subgraphs of $C$.
        \item The vertex set of $\bigcup_{i=1}^N{C_i}$ is the 
            vertex set of $C$.
        \item A leaf node is a trivial subgraph.
       \item A root node is a maximal rigid subgraph of $G$
    \end{enumerate}

@@

    A DR-plan is \textbf{complete} if it satisfies an additional rule: 
        \item For a node, $C$, that is  a non-trivial graph, its $N$
            children, $C_1, \ldots, C_N$, are all of the 
            rigid vertex-maximal proper subgraphs of $C$.
    This makes Rule 2 implicit, and in fact, $C$ contains all the edges in
    the union of its children as well. The complete DR-plan is unique.

    A DR-plan is \textbf{optimal} if it minimizes the maximum fan-in over all nodes in the tree. It is not necessarily unique.
\end{definition}

@@

\subsection*{Importance of an Optimal DR-plan}
The use of a DR-plan in designing especially self-similar qusecs is
immediately clear, as shown in Figures \ref{XX} and \ref{XX}.
In fact, knowing an optimal DR-plan is crucial for design and analysis of 
even non-self-similar qusecs 
and their realizations with various desirable properties.

>>>>

\subsection{Basics}

In this paper, a {\em geometric constraint system} is a multivariate polynomial 
(usually bilinear or quadratic) system $G(x,\delta)=0$, representing constraints with parameters $\delta$ 
between geometric primitives  in $\mathbb{R}^2$ represented collectively as $x\in \mathbb{R}^n$.   
%
When the type of constraint (system) is fixed, the system is simply represented as $(G,\delta)$, where 
$G$ is the underlying constraint (hyper)graph $G = (V,E)$ 
with the vertices $V$ representing the geometric primitives in 
$\mathbb{R}^2$ and (hyper)edges
$E$ representing the constraints, each with an associated 
parameter $\delta$.
%
For example, a {\em bar-joint system or linkage} $(G,\delta)$, is a graph
$G=(V,E)$  
with fixed length bars as edges, i.e. $\delta: E \rightarrow \mathbb{R}$;
this represents the distance constraint system $\| x_u -x_v \|_2 =
\delta_{u,v}$ for  $(u,v) \in E$, where $x_u \in \mathbb{R}^2$ represents the coordinates of
$u\in V$.

\noindent {\bf Note:} Geometric constraint systems can also have inequalities, 
where the parameters in $\delta$ are small intervals of values rather than exact values.

@@

%XXXXXXXXXsomething has to be said 
%along with asimow roth (infinitesimal rigidity 
%implies rigidity)
%When the rigidity matrix has appropriate rank, the realizations or solutions 
%of the corresponding constraint system are generically isolated and zero-dimensional 
%if the constraint system has equalities and exact values for the parameters $\delta$.
%If the constraint system has inequalities, i.e, if the $\delta$ lie in an interval,
%this claim is approximate: the solutions are isolated small, full-dimensional  connected components.

One important property of generic constraint system or 
(hyper)graph %that depends on the appropriate rigidity matrix of indeterminates
is \emph{rigidity} ({\bf Note:} we refer to these as
properties of the constraint system or as properties of the underlying
(hyper)graph interchangeably),
i.e., the realizations or solutions of the corresponding constraint system being generically isolated and zero-dimensional. %(if the constraint system has equalities and exact values for the parameters $\delta$). 
The result by Asimow and Roth~\cite{???} shows a constraint (hyper)graph is rigid
if it is generically {\em infinitesimally rigid}, i.e.\ the number of independent rows of its appropriate rigidity matrix is at least the number of columns less the number of rigid body motions, which is 3 for bar-joint systems. 
%This condition is called generic {\em infinitesimal rigidity}.

\noindent {\bf Note:} If the constraint system has inequalities, i.e, if the $\delta$ lie in an interval,
the definition of rigidity is approximate: the solutions are isolated small, full-dimensional  connected components.
% generic rigidity, i.e.\ there exist at most finitely many solutions to the system at generic points,  %to the algebraic system

Other generic constraint system or (hyper)graph properties are the following: %({\bf Note:} we refer to these as
%properties of the constraint system or as properties of the underlying
%(hyper)graph interchangeably).
%
A constraint (hyper)graph $G$ is {\em independent} if its appropriate rigidity matrix of indeterminates  
has independent rows
(i.e, the determinant of some square submatrix is not
identically zero). 
%It is {\em rigid} if the number of independent rows of the rigidity matrix is
%at least the number of columns less the number of rigid body motions, 
%which is 3 for distance constraint systems. 
It is {\em isostatic (minimally rigid, well-constrained)} 
if it is both rigid and independent. 
%if the number of generically independent rows or the rank of the appropriate rigidity matrix is maximal.
%For example, the maximal rank of a bar-joint rigidity matrix is $2|V| - 3$, 
%where $3$ is the number of rotational and translational degrees-of-freedom of a rigid body in $\mathbb{R}^2$.
%The graph $G$ is {\em rigid} if there exists some spanning subgraph $S\subseteq G$ such that $S$ is well-constrained. 
It is {\em flexible} if it is not rigid, {\em under-constrained} if it is independent and not rigid, or {\em over-constrained} if it is not independent.

Defining the independence of a subset of edges $E' \subseteq E$ 
to be the independence of corresponding rows in the appropriate rigidity matrix of indeterminates,
we obtain the {\em rigidity matroid} of a constraint (hyper)graph $G = (V,E)$. 
%The 2-dimensional {\em rigidity matroid} of a constraint (hyper)graph $G = (V,E)$ is a linear matroid  on $E$, 
%where a subset of edges $E' \subseteq E$ is {\em independent} in the matroid, 
%if the set of corresponding rows in the appropriate rigidity matrix of indeterminates are linearly independent. 
There are various results on combinatorial characterization of independence and rigidity
for different types of  rigidity matroid. 
For bar-joint rigidity matroid, 
the famous Laman's theorem~\cite{???} states that the underlying graph
is minimally rigid if and only if $|E| = 2|V|-3$ and $|E'| \le 2|V'|-3$ for every induced subgraph with at least 2 vertices. 
The result by Lovasz and Yemini~\cite{???} shows that all 6-vertex-connected graphs are rigid in the plane. 
For bar-body rigidity matroid,
Tay~\cite{???} proved that
the underlying multigraph is minimally rigid if and only if it can be decomposed as $3$ edge disjoint spanning trees. 
White and Whiteley~\cite{???} gave the same characterization using a different technique to study the algebraic - geomtric conditions called pure condition.
Lee, Streinu and Theran~\cite{???} defined the {\em $(k,l)$-sparsity matroid},
where a hypergraph $G$ is called {\em $(k,l)$-sparse} if $|E'| \le k|V'| - l$ for any induced subgraph $(V',E')$ with at least 2 vertices, 
and {\em $(k,l)$-tight} if is $(k,l)$-sparse and $|E| = k|V| - l$. 
In general, given a $d$-uniform hypergraph, a $(k,l)$-sparsity
condition is matroidal as long as $l \le dk-1$.

In this paper, a {\em qusec} is any {\em independent} geometric 
constraint system of one of the 3 types mentioned above.
In the remainder of this section and the next two sections~\ref{sec:DRP} and \ref{sec:recomb} 
we only consider bar-joint qusecs and graphs. 
Relevant formal analogies  for the other 2 types of qusecs and (hyper)graphs 
are given in the subsequent 2 sections on Applications.

%XXXNow comes the definition of stress vector and flex vector. 
%NOTE: something has to be said about what flexes and stresses mean

%TODO
Given a bar-joint graph, a vector (of indeterminates) in the right null space of its rigidity matrix
is called a {\em flex vector}. 
It has 2 entries per vertex and represents the internal motions of the system
(i.e.\ modulo rigid body motions).
For rigid graphs, all flex vectors are identically zero. 
%
A vector (of indeterminates) in the left null space of the rigidity matrix
is called a {\em stress vector}.
It has one entry per edge and represents a {\em self stress} of the system.
For independent graphs, all stress vectors are identically zero. 
%
An {\em external stress} $t$ is a $1\times 2|V|$ vector  (of indeterminates)
which specifies a 2D vector acting at each vertex
such that the stress balance equation $sR = t$ holds, where $s$ is a vector
of internal stresses, one for each bar, and $R$ is the rigidity matrix.

\subsection{Decomposition-Recombination (DR) Plans}

@@

A {\em trivial} graph is   a single vertex.
\begin{definition}\label{def:drp}
    The  {\em decomposition-recombination \mbox{(DR-)} plan} of graph $G$,
    \cite{XX}
    $DRP(G)$, is defined as a forest that has the following properties:
    \begin{enumerate}
        \item Each  node represents/contains/is  a rigid vertex-induced subgraph of $G$
        %\item For a node, $C$, that is  a non-trivial graph, its $N$
        %    children, $C_1, \ldots, C_N$, are rigid vertex-maximal proper subgraphs of $C$.
        \item The vertex set of $\bigcup_{i=1}^N{C_i}$ is the 
            vertex set of $C$.
        \item A leaf node is a trivial subgraph.
       \item A root node is a maximal rigid subgraph of $G$
    \end{enumerate}

@@

    A DR-plan is {\em complete} if it satisfies an additional rule: 
        \item For a node, $C$, that is  a non-trivial graph, its $N$
            children, $C_1, \ldots, C_N$, are all of the 
            rigid vertex-maximal proper subgraphs of $C$.
    This makes Rule 2 implicit, and in fact, $C$ contains all the edges in
    the union of its children as well. The complete DR-plan is unique.

    A DR-plan is {\em optimal} if it minimizes the maximum fan-in over all nodes in the tree. It is not necessarily unique.
\end{definition}

@@

\subsection{Importance of an Optimal DR-plan}
The use of a DR-plan in designing especially self-similar qusecs is
immediately clear, as shown in Figures \ref{XX} and \ref{XX}.
In fact, knowing an optimal DR-plan is crucial for design and analysis of 
even non-self-similar qusecs 
and their realizations with various desirable properties.

### 2014-12-17 doc/GMP2015/s04_recombination.tex
Updated preliminary. Added not to recombination. Modified pinned line-incidence according to prelimary and updated DRP section.

In the following, we formulate the problem of {\em optimal
modification},
of the indecomposable
algebraic system $I$ into a decomposable system with a small DR-plan (low
algebraic complexity).
Leveraging recent results on {\em Cayley
configuration spaces}, our approach to the optimal modification problem
achieves the following: {\em (a) Small DR-plan.} We obtain a  parameterized family of systems
$I_{\lambda_F}$ -  one for each value $\lambda_F$ for the parameters $F$ ,  all of which have
small DR-plans. Thus, given a value $v$ for $\lambda_F$, the system $I_v$ can
potentially be solved or realized easily once the orientation type of the solution
is known  (when the DR-plan size is small enough).
{\em (b) Solution preservation.} Moreover, the union of solution spaces of the systems in the family
$I_{\lambda_F}$
is guaranteed to contain
all of $I$'s solutions. {\em (c) Efficient search.} Finally, this union can be projected into the
so-called {\em Cayley} or distance parameter space $\lambda_F$ that is convex
or otherwise easy to traverse in order to search for $I$'s solution
(or connected component)
of the desired orientation type.
For the case when the modification is bounded, this approach provides an efficient
algorithm for recombination. We first define the decision version of the
problem of optimal modification
for decomposition. The standard optimization versions are straightforward.

@@

\caption{
(\ref{fig:omd_k33_example:a}) The $K3,3$ with two labeled edges, $e_1$ and $e_2$. (\ref{fig:omd_k33_example:b}) The $K3,3$ with $e_1$ and $e_2$ removed (dashed lines) and rearranged to illustrate that it is now a partial 2-tree. (\ref{fig:omd_k33_example:c}) The $K3,3$ with $\{e_1,e_2\}$ removed and $\{f_1,f_2\}$ (bold lines) added to make a 2-tree, showing that the $K3,3$ is at least $OMD_2$. (\ref{fig:omd_k33_example:d}) The $K3,3$ with only $e_2$ removed (dashed line). (\ref{fig:omd_k33_example:e}) The $K3,3$ with $e_2$ removed and $f_3$ (bold line) added to make a low Cayley complexity graph, showing that the $K3,3$ is $OMD_1$.
}
\label{fig:omd_k33_example}
\end{figure*}

\medskip\noindent{\bf Example}.
A graph $G=K_{3,3}$  cannot be decomposed into any nontrivial isostatic
graphs, i.e, its DR-plan has a root and 9
children corresponding to the 9 edges. Solving or recombining the system
$(G,\delta)$
corresponding to the root of
this DR-plan involves solving a quadratic system with 8 equations and variables,
Instead of simultaneously solving this system, we could instead use the fact
that $G=K_{3,3}$ is in OMD$_1$: remove the
edge $e$ in Figure \ref{fig:omd_k33_example} to give a partial 2-tree $G'$.
Now add the nonedge $f$ to give a 2-tree $H$ with a DR-plan of size 2.
The Cayley configuration space $\Phi_f(G', \delta_{E\setminus e})$
is a single interval of attainable length values $\lambda_F$ for
the edge $f$.
When $\delta$ is generic, i.e, does not admit collinearities or coincidences
in the realizations of $(G,\delta)$,
the realization space of $(H, <\delta_{E\setminus e}, \lambda_f>)$ has 16
solutions $q_p^{\lambda_f}$ (modulo orientation preserving isometries),
with distinct orientation types $\sigma_p$,
(two orientation choices for each of the 4 triangles)
that can be obtained by solving a sequence of 4 single quadratics in 1
variable (DR-plan of size 2).
By subdivided
binary search
in the interval $\lambda_f \in
\Phi_f(G', \delta_{E\setminus e})$, the desired
solution $p$  of $(G,\delta)$ is found when the length of the nonedge
$e$ in  the realization
$q^p_{\lambda_f}$
is $\delta_e$.
%
\subsection{Optimized Modification by Enlarging the class of Reduced
Graphs}
\label{tdecomp}
It is possible in principle to decrease $k$ for a OMD$_k$ graph
(i.e, the number of edges to be removed
to ensure an isostatic completion that is decomposable with a small DR-plan) by
considering reduced graphs $G'$ (and modified graphs $H$) that come from a larger class than
partial 2-trees but nevertheless have convex Cayley configuration spaces at
least when the realization space is restricted to a sufficiently
comprehensive orientation type.
In particular, the so-called {\em tree-decomposable graphs of low Cayley
complexity} \cite{XX} include the partial 2-trees and many others that are
not partial 2-trees. These too result in DR-plans of size 2 or 3, putting $G$ in the class
OMD$_k$ and thus meeting Criteria (a) and (b). The Criterion (c) is met
- for example when $k=1$ -  because 1 dof Cayley configuration spaces of
linkages based on such graphs $G'$
are known to be single intervals
when a comprehensive orientation type $\sigma_p$ of the sought solution $p$ is given.
In addition, the bounds of these intervals are of low algebraic
complexity.
More precisely, the bounds  can themselves be computed using a DR-plan of size 2 or 3,
i.e, the computation of these bounds is tree-decomposable. Alternatively,
the bounds are in a simple quadratic or radically solvable extension field over the rationals,
or can be computed by solving a
triangularized system of quadratics.

>>>>

In the following, we formulate the problem of {\em optimal
modification},
of the indecomposable
algebraic system $I$ into a decomposable system with a small DR-plan (low
algebraic complexity).
Leveraging recent results on {\em Cayley
configuration spaces}, our approach to the optimal modification problem
achieves the following: {\em (a) Small DR-plan.} We obtain a  parameterized family of systems
$I_{\lambda_F}$ -  one for each value $\lambda_F$ for the parameters $F$,  all of which have
small DR-plans. Thus, given a value $v$ for $\lambda_F$, the system $I_v$ can
potentially be solved or realized easily once the orientation type of the solution
is known  (when the DR-plan size is small enough).
{\em (b) Solution preservation.} Moreover, the union of solution spaces of the systems in the family
$I_{\lambda_F}$
is guaranteed to contain
all of $I$'s solutions. {\em (c) Efficient search.} Finally, this union can be projected into the
so-called {\em Cayley} or distance parameter space $\lambda_F$ that is convex
or otherwise easy to traverse in order to search for $I$'s solution
(or connected component)
of the desired orientation type.
For the case when the modification is bounded, this approach provides an efficient
algorithm for recombination. We first define the decision version of the
problem of optimal modification
for decomposition. The standard optimization versions are straightforward.

@@

\caption{
(\ref{fig:omd_k33_example:a}) The $K_{3,3}$ with two labeled edges, $e_1$ and $e_2$. (\ref{fig:omd_k33_example:b}) The $K_{3,3}$ with $e_1$ and $e_2$ removed (dashed lines) and rearranged to illustrate that it is now a partial 2-tree. (\ref{fig:omd_k33_example:c}) The $K_{3,3}$ with $\{e_1,e_2\}$ removed and $\{f_1,f_2\}$ (bold lines) added to make a 2-tree, showing that the $K_{3,3}$ is at least $OMD_2$. (\ref{fig:omd_k33_example:d}) The $K_{3,3}$ with only $e_2$ removed (dashed line). (\ref{fig:omd_k33_example:e}) The $K_{3,3}$ with $e_2$ removed and $f_3$ (bold line) added to make a low Cayley complexity graph, showing that the $K_{3,3}$ is $OMD_1$.
}
\label{fig:omd_k33_example}
\end{figure*}

\noindent{\bf Note:} A huge advantage of the convex Cayley method
is that it is completely unaffected when $\delta$ are intervals rather than exact values~\cite{GaoSitharam}.
%cite my paper with Heping!!!

\medskip\noindent{\bf Example}.
A graph $G=K_{3,3}$  cannot be decomposed into any nontrivial isostatic
graphs, i.e, its DR-plan has a root and 9
children corresponding to the 9 edges. Solving or recombining the system
$(G,\delta)$
corresponding to the root of
this DR-plan involves solving a quadratic system with 8 equations and variables.
Instead of simultaneously solving this system, we could instead use the fact
that $G=K_{3,3}$ is in OMD$_1$: remove the
edge $e$ in Figure \ref{fig:omd_k33_example} to give a partial 2-tree $G'$.
Now add the nonedge $f$ to give a 2-tree $H$ with a DR-plan of size 2.
The Cayley configuration space $\Phi_f(G', \delta_{E\setminus e})$
is a single interval of attainable length values $\lambda_F$ for
the edge $f$.
When $\delta$ is generic, i.e, does not admit collinearities or coincidences
in the realizations of $(G,\delta)$,
the realization space of $(H, <\delta_{E\setminus e}, \lambda_f>)$ has 16
solutions $q^p_{\lambda_f}$ (modulo orientation preserving isometries),
with distinct orientation types $\sigma_p$ 
(two orientation choices for each of the 4 triangles)
that can be obtained by solving a sequence of 4 single quadratics in 1
variable (DR-plan of size 2).
By subdivided
binary search
in the interval $\lambda_f \in
\Phi_f(G', \delta_{E\setminus e})$, the desired
solution $p$  of $(G,\delta)$ is found when the length of the nonedge
$e$ in  the realization
$q^p_{\lambda_f}$
is $\delta_e$.
%
\subsection{Optimized Modification by Enlarging the class of Reduced
Graphs}
\label{tdecomp}
It is possible in principle to decrease $k$ for a OMD$_k$ graph
(i.e, the number of edges to be removed
to ensure an isostatic completion that is decomposable with a small DR-plan) by
considering reduced graphs $G'$ (and modified graphs $H$) that come from a larger class than
partial 2-trees but nevertheless have convex Cayley configuration spaces at
least when the realization space is restricted to a sufficiently
comprehensive orientation type.
In particular, the so-called {\em tree-decomposable graphs of low Cayley
complexity} \cite{XX} include the partial 2-trees and many others that are
not partial 2-trees. These too result in DR-plans of size 2 or 3, putting $G$ in the class
OMD$_k$ and thus meeting Criteria (a) and (b). The Criterion (c) is met
- for example when $k=1$ -  because 1 dof Cayley configuration spaces of
linkages based on such graphs $G'$
are known to be single intervals
when a comprehensive orientation type $\sigma_p$ of the sought solution $p$ is given.
In addition, the bounds of these intervals are of low algebraic
complexity.
More precisely, the bounds  can themselves be computed using a DR-plan of size 2 or 3,
i.e, the computation of these bounds is tree-decomposable. Alternatively,
the bounds are in a simple quadratic or radically solvable extension field over the rationals,
or can be computed by solving a
triangularized system of quadratics.
